Synthesizing Symmetric Lenses

### 2018-07-03 papers/SymmetricLenses/main.tex
Some more comments

\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{amsmath}
\usepackage{nccmath}
\usepackage{mathtools}
\usepackage{bussproofs}
\usepackage{varwidth}
\usepackage{amsthm}
\usepackage{csvsimple}
\usepackage{thmtools,thm-restate}
\usepackage{changepage}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage{enumitem}
\usepackage{multirow,bigdelim}
\usepackage{multicol}
\usepackage{siunitx}
\usepackage{listings}
\usepackage{letltxmacro}
\usepackage{sansmath}
\usepackage{url}
\usepackage{flushend}
\usepackage{microtype}
\usepackage[utf8]{inputenc}
\usepackage{mathpartir}
\usepackage{empheq}
\usepackage{array}
\usepackage{pgfplots}
\usepackage{stmaryrd}
\usepackage{courier}
\usepackage{qtree}
\usepackage[normalem]{ulem}
\usepackage{relsize}
\usepackage{tikz}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{textcomp}
\usepackage{tabularx}
\usepackage{stackengine}
\usepackage{caption}
\usepackage{wrapfig}
\usepackage{remreset}
\usepackage{tabulary}
\usepackage{xspace}

@@

%Consider the trivial lens
%\lstinline{disc(emp_salary,emp_row,rep(emp_salary),rep(emp_row))}. As this is
%the disconnect lens, the cost of each side is the entropy of the data format, so
%the total cost is $\EntropyOf{\lstinline{emp_salary}} +
%\EntropyOf{\lstinline{emp_row}}$. \lstinline{emp_salary} contains strictly more
%imformation than \EntropyOf{\lstinline{": " . salary}}, and \lstinline{emp_row}
%contains strictly more information than \EntropyOf{\lstinline{": " . salary}},
%so by our metric \lstinline{single_emp_map} is preferable to the disconnect
%lens.
Consider the trivial lens
\lstinline{disc(emp_salary,emp_row,rep(emp_salary),rep(emp_row))}. As this is
the disconnect lens, the cost of each side is the entropy of the data format, so
the total cost is $\EntropyOf{\lstinline{emp_salary} } +
\EntropyOf{\lstinline{emp_row}}$. \lstinline{emp_salary} contains strictly more
imformation than $\EntropyOf{\lstinline{": " . salary}}$, and
\lstinline{emp_row} contains strictly more information than
\EntropyOf{\lstinline{": " . salary}}, so by our metric
\lstinline{single_emp_map} is preferable to the disconnect lens.
% end overview

@@

    $\ProbabilityOf{\PRegexStar{\Regex}{\Probability}}{\String}$
    & =
    & $\Sigma_{\String = \String_1 \ldots \String_n}\Probability^n*(1-\Probability)*\Pi_{i=1}^n\ProbabilityOf{\Regex}{\String_i}$\\
  \end{tabular}
\end{center}
In the stochastic regular expression
$\PRegexOr{\Regex_1}{\Regex_2}{\Probability}$, the annotation \Probability
represents the likelihood the string is in $\Regex_1$. In the stochastic regular
expression $\PRegexStar{\Regex}{\Probability}$, the annotation \Probability
how much more likely having a string involving more than $n$ iterations is than
having a string involving $n$ iterations.

@@

\begin{center}
  \begin{tabular}{@{}r@{\hspace{1em}}c@{\hspace{1em}}l@{}r@{}}
    \PRegexOr{\Regex}{\emptyset}{1} & $\SSREquiv$ & \Regex{} & \OrIdentityRule{} \\
    $\RegexConcat{\Regex}{\emptyset}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionRightRule{} \\
    $\RegexConcat{\emptyset}{\Regex}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionLeftRule{} \\
    \RegexConcat{(\RegexConcat{\Regex{}}{\Regex'})}{\Regex''} & $\SSREquiv$ & \RegexConcat{\Regex{}}{(\RegexConcat{\Regex'}{\Regex''})} & \ConcatAssocRule{}  \\
    \PRegexOr{(\PRegexOr{\Regex}{\Regex'}{\Probability_1})}{\Regex''}{\Probability_2} & $\SSREquiv$ & \PRegexOr{\Regex}{(\PRegexOr{\Regex'}{\Regex''}{(1-\Probability_1)*\Probability_2})}{\Probability_1*\Probability_2} & \OrAssociativityRule{}  \\
    \PRegexOr{\Regex{}}{\RegexAlt{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\RegexAlt{}}{\Regex{}}{1-\Probability} & \OrCommutativityRule{}\\
    \RegexConcat{\Regex{}}{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}}{\Regex{}'})}{(\RegexConcat{\Regex{}}{\Regex{}''})}{\Probability} & \DistributivityLeftRule{} \\
    \RegexConcat{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})}{\Regex{}} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}'}{\Regex{}})}{(\RegexConcat{\Regex{}''}{\Regex{}})}{\Probability} & \DistributivityRightRule{} \\
    \RegexConcat{\EmptyString{}}{\Regex{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityLeftRule{} \\
    \RegexConcat{\Regex{}}{\EmptyString{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityRightRule{} \\
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{\Probability} & \UnrollstarLeftRule{} \\
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{{\Regex{}}}{\Probability}}{\Regex{}})}{\Probability} & \UnrollstarRightRule{} 
  \end{tabular}
\end{center}

@@

\subsection{Lens Costs}
Our cost metric is based on the expected amount of information required to
recover a string in one data format from the other. We use the function
$\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ to calculate the expected amount
of information required to recover a string in $\Regex$ from a string in
$\RegexAlt$, synchronized by $\Lens$. We use the function
$\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ to calculate the expected amount
of information required to recover a string in $\RegexAlt$ from a string in
$\Regex$, synchronized by $\Lens$.
\begin{center}
  \begin{tabular}{rcl}
    $\LEntropyOf{\Regex \Given \IdentityLensOf{\Regex}, \Regex}$
    & =
    & 0\\

    $\LEntropyOf{\Regex \Given \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}, \RegexAlt}$
    & =
    & \EntropyOf{\Regex}\\

    $\LEntropyOf{\PRegexStar{\Regex}{\Probability} \Given \IterateLensOf{\Lens}^, \PRegexStar{\RegexAlt}{\ProbabilityAlt}}$
    & =
    & $\frac{\ProbabilityAlt}{1-\ProbabilityAlt}(\LEntropyOf{\Regex \Given \Lens, \RegexAlt})$\\

    $\LEntropyOf{\Regex_1 \Concat \Regex_2 \Given \ConcatLensOf{\Lens_1}{\Lens_2}, \RegexAlt_1 \Concat \RegexAlt_2}$
    & =
    & $\LEntropyOf{\Regex_1 \Given \Lens_1^\leftarrow, \RegexAlt_1} + \LEntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt_2}$\\

    $\LEntropyOf{\Regex_1 \Concat \Regex_2 \Given \SwapLensOf{\Lens_1}{\Lens_2}, \RegexAlt_2 \Concat \RegexAlt_1}$
    & =
    & $\LEntropyOf{\Regex_1 \Given \Lens_1^\leftarrow, \RegexAlt_1} + \LEntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt_2}$\\

    $\LEntropyOf{\PRegexOr{\Regex_1}{\Regex_2}{\Probability} \Given \OrLensOf{\Lens_1}{\Lens_2}, \PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt}}$
    & =
    & $\ProbabilityAlt*(\LEntropyOf{\Regex_1 \Given \Lens_1, \RegexAlt_1}) + (1-\ProbabilityAlt)*(\LEntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt_2})$\\

    $\LEntropyOf{\PRegexOr{\Regex_1}{\Regex_2}{\Probability} \Given \OrLensOf{\Lens_1}{\Lens_2}, \PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt}}$
    & =
    & $\ProbabilityAlt*(\LEntropyOf{\Regex_1 \Given \Lens_1, \RegexAlt_1}) + (1-\ProbabilityAlt)*(\LEntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt_2})$\\

    $\LEntropyOf{\Regex \Given \MergeROf{\Lens_1}{\Lens_2}, \PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt}}$
    & =
    & $\ProbabilityAlt*(\LEntropyOf{\Regex \Given \Lens_1, \RegexAlt_1}) + (1-\ProbabilityAlt)*(\LEntropyOf{\Regex \Given \Lens_2, \RegexAlt_2})$\\

    $\LEntropyOf{\PRegexOr{\Regex_1}{\Regex_2}{\Probability} \Given \MergeLOf{\Lens_1}{\Lens_2}, \RegexAlt}$
    & =
    & $\Probability*(\LEntropyOf{\Regex_1 \Given \Lens_1, \RegexAlt} - \Log_2(\Probability)) + $\\
    &
    & $(1-\Probability)*(\LEntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt} - \Log_2(1 - \Probability))$\\

    $\LEntropyOf{\RegexAlt \Given \InvertOf{\Lens}, \Regex}$
    & =
    & $\REntropyOf{\RegexAlt \Given \Lens, \Regex}$\\
  \end{tabular}
\end{center}
$\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ is defined symmetrically.  These
functions calculate the expected number of bits to recover one data format from
a synchronized string in the other format.
The cost of a lens, $c(\Lens \OfType \Regex \Leftrightarrow \RegexAlt) =
\LEntropyOf{\Regex \Given \Lens, \RegexAlt} + \REntropyOf{\RegexAlt \Given
  \Lens, \Regex}$ is the
sum of recovering the left format from the right, and the right from the left.
We have proven theorems demonstrating the calculated entropy corresponds to the
actual conditional entropy to recover the data.\bcp{Again, they should be
  included.}

\begin{theorem}
  Let $\Lens \OfType \Regex \Leftrightarrow \RegexAlt$, where $\Lens$ does not
  include composition, $\Regex$ and $\RegexAlt$ and unambiguous, and neither
  $\Regex$ nor $\RegexAlt$ contain any empty subcomponents.
  \begin{enumerate}
  \item $\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ calculates the expected
    information content of $\SetOf{s \SuchThat s \in \LanguageOf{\Regex}}$,
    given $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt} \BooleanAnd
      \Lens.\PutLOf{t}{s} = s}$
  \item $\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ calculates the expected
    information content of $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt}}$,
    given $\SetOf{s \SuchThat s \in \LanguageOf{\Regex} \BooleanAnd
      \Lens.\PutROf{s}{t} = t}$
  \end{enumerate}
\end{theorem}

@@

As these DNF regular expressions are \emph{stochastic}, they are annotated with
probabilities to express a probability distribution, in addition to a language.
\begin{center}
  \begin{tabular}{rcl}
    $\ProbabilityOf{\PRegexStar{\DNFRegex}{\Probability}}{\String}$
    & =
    & $\Sigma_{\String = \String_1 \ldots \String_n}\Probability^n*(1-\Probability)*\Pi_{i=1}^n\ProbabilityOf{\DNFRegex}{\String_i}$\\

@@

    $\EntropyOf{\SequenceOf{\String_0 \SeqSep \Atom_1 \SeqSep \ldots \SeqSep \Atom_n \SeqSep \String_n}}{\String'}$
    & =
    & $\Sigma_{i = 1}^n\EntropyOf{\Atom_i}$ \\

@@

Much like we defined a syntactic means to find the expected number of bits to
recover string from a synchronized string in another format for simple symmetric
string lenses, we do the same for symmetric DNF lenses.  Unlike simple symmetric
string lenses, DNF lenses do not have composition, so the entropy can be defined
over all terms.  $\LEntropyOf{\DNFRegex \Given \SDNFLens, \DNFRegexAlt}$ is the
expected number of bits required to recover a string in $\DNFRegex$ from a
synchronized string in $\DNFRegexAlt$.

\[
  \begin{array}{rl}
    & \LEntropyOf{\DNFOf{\Sequence_1 \DNFSep \ldots \DNFSep \Sequence_n}\\
    & \hspace*{1.21em}\Given
      (\SDNFLensOf{(i_1,j_1,\SSQLens_1)\DNFLSep
      \ldots\DNFLSep
      (i_p,j_p,\SSQLens_p)}
      ,\ListOf{c_1;\ldots;c_q}
      ,\ListOf{d_1;\ldots;d_r})\\
    & \hspace*{1.21em},~\DNFOf{\SequenceAlt_1 \DNFSep \ldots \DNFSep \SequenceAlt_m}}\\
    = & \frac{\Sigma_{j=1}^m(H_j)}{m}\\
    & \text{where } H_j = \frac{\Sigma_{k\SuchThat j_k = j}
      \LEntropyOf{\Sequence_{i_k}\Given\SequenceLens_k,\SequenceAlt_j}}{\SizeOf{\SetOf{k\SuchThat
      j_k = j}}} + log_2\SizeOf{\SetOf{k\SuchThat j_k = j}}
  \end{array}
\]

@@

\[
  \begin{array}{rl}
      & \LEntropyOf{
        \SequenceOf{\String_0,\Atom_1,\ldots,\Atom_n,\String_n}\\
      & \hspace*{1em}\Given
        (\SSQLensOf{(i_1,j_1,\SAtomLens_1)\SeqLSep
        \ldots\SeqLSep
        (i_p,j_p,\SAtomLens_p)}
        ,\ListOf{(k_1,\String_1);\ldots;(k_q,\String_q)}
        ,\ListOf{(l_1,\StringAlt_1);\ldots;(l_r,\StringAlt_r)})\\
      & \hspace*{1.21em},~
        \SequenceOf{\StringAlt_0,\AtomAlt_1,\ldots,\AtomAlt_m,\StringAlt_m}}\\
      =
      & \Sigma_{x=1}^p\LEntropyOf{\Atom_{i_x} \Given \AtomAlt_{i_x},\AtomLens_x} +
        \Sigma_{x=1}^q\LEntropyOf{\Atom_{k_x}}
  \end{array}
\]

@@

\[\begin{array}{rl}
      & \LEntropyOf{\PRegexStar{\DNFRegex}{\Probability} \Given \StarOf{\DNFRegexAlt},\StarOf{\SDNFLens}}\\
      =\\
      & 4*\LEntropyOf{\DNFRegex\Given\DNFRegexAlt,\SDNFLens}
  \end{array}
\]
The expected number of bits required to recover a string in $\DNFRegexAlt$ from
a synchronized string in $\DNFRegex$, $\REntropyOf{\DNFRegexAlt \Given
  \SDNFLens, \DNFRegex}$ is defined symmetrically.

@@

\paragraph*{Compositional Synthesis} Compositional synthesis allows us to use
previously defined lenses - whether they are previously defined by users or by
synthesis. We alter \GreedySynth such that two regular expression atoms can be
mapped to each other either if they are the same regular expression and have
examples agreeing with the examples, or if there is a lens which is a
composition of predefined lenses between them, where the lens alters the
examples in a desired way.

\paragraph*{\RXSearch Permissiveness Customization}
While termination heuristic works for many situations: there are situations
where a large number of expansions must be performed to find the correct lens.
While users could prevent incorrect lenses with a large number of examples, we
find it simpler to let the users specify that \RXSearch should be more
restrictive in which lenses it accepts. If the user's desired transformation has
a particularly high cost, the algorithm may spend a long time searching in vain
for lenses with lower cost.  By tuning \RXSearch to me more permissive in the
lenses it accepts, this search can be short-circuited.
% end synthesis

@@

We have extended Optician with our symmetric synthesis algorithm. Users can
write synthesis tasks alongside lens combinators, use those synthesis results in
the manual writing of new lenses, and use previously defined lenses in the
synthesis of new lenses.

In this work, we aim to answer two primary research questions:
\begin{enumerate}
\item Is our synthesis procedure efficient enough to be used in everyday
  development?  Can it be used in place of the existing bijective lens synthesis
  algorithm integrated with Boomerang?
\item Are our heuristics useful for reducing the specifications users have to
  provide.
\end{enumerate}

>>>>

\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{amsmath}
\usepackage{nccmath}
\usepackage{mathtools}
\usepackage{bussproofs}
\usepackage{varwidth}
\usepackage{amsthm}
\usepackage{csvsimple}
\usepackage{thmtools,thm-restate}
\usepackage{changepage}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage{enumitem}
\usepackage{multirow,bigdelim}
\usepackage{multicol}
\usepackage{siunitx}
\usepackage{listings}
\usepackage{letltxmacro}
\usepackage{sansmath}
\usepackage{url}
\usepackage{flushend}
\usepackage{microtype}
\usepackage[utf8]{inputenc}
\usepackage{mathpartir}
\usepackage{empheq}
\usepackage{array}
\usepackage{pgfplots}
\usepackage{stmaryrd}
\usepackage{courier}
\usepackage{qtree}
\usepackage[normalem]{ulem}
\usepackage{relsize}
\usepackage{tikz}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{textcomp}
\usepackage{tabularx}
\usepackage{stackengine}
\usepackage{caption}
\usepackage{wrapfig}
\usepackage{remreset}
\usepackage{tabulary}
\usepackage{xspace}
\usepackage{bbm}

\newtheorem*{theorem*}{Theorem}

@@

%Consider the trivial lens
%\lstinline{disc(emp_salary,emp_row,rep(emp_salary),rep(emp_row))}. As this is
%the disconnect lens, the cost of each side is the entropy of the data format, so
%the total cost is $\EntropyOf{\lstinline{emp_salary}} +
%\EntropyOf{\lstinline{emp_row}}$. \lstinline{emp_salary} contains strictly more
%imformation than \EntropyOf{\lstinline{": " . salary}}, and \lstinline{emp_row}
%contains strictly more information than \EntropyOf{\lstinline{": " . salary}},
%so by our metric \lstinline{single_emp_map} is preferable to the disconnect
%lens.
Consider the trivial lens
\lstinline{disc(emp_salary,emp_row,rep(emp_salary),rep(emp_row))}. As this is
the disconnect lens, the cost of each side is the entropy of the data format, so
the total cost is $\EntropyOf{\lstinline{emp_salary} } +
\EntropyOf{\lstinline{emp_row}}$. \lstinline{emp_salary} contains strictly more
information than $\EntropyOf{\lstinline{": " . salary}}$, and
\lstinline{emp_row} contains strictly more information than
\EntropyOf{\lstinline{": " . salary}}, so by our metric
\lstinline{single_emp_map} is preferable to the disconnect lens.
% end overview

@@

    $\ProbabilityOf{\PRegexStar{\Regex}{\Probability}}{\String}$
    & =
    & $\Sigma_n \Sigma_{\String = \String_1 \ldots \String_n}\Probability^n*(1-\Probability)*\Pi_{i=1}^n\ProbabilityOf{\Regex}{\String_i}$\\
  \end{tabular}
\end{center}
In the stochastic regular expression
$\PRegexOr{\Regex_1}{\Regex_2}{\Probability}$, the annotation \Probability
represents the likelihood the string is in $\Regex_1$. In the stochastic regular
expression $\PRegexStar{\Regex}{\Probability}$, the annotation \Probability
how much more likely having a string involving more than $n$ iterations is than
having a string involving $n$ iterations.

@@

\begin{center}
  \begin{tabular}{@{}r@{\hspace{1em}}c@{\hspace{1em}}l@{}r@{}}
    \PRegexOr{\Regex}{\emptyset}{1} & $\SSREquiv$ & \Regex{} & \OrIdentityRule{} \\
    $\RegexConcat{\Regex}{\emptyset}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionRightRule{} \\
    $\RegexConcat{\emptyset}{\Regex}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionLeftRule{} \\
    \RegexConcat{(\RegexConcat{\Regex{}}{\Regex'})}{\Regex''} & $\SSREquiv$ & \RegexConcat{\Regex{}}{(\RegexConcat{\Regex'}{\Regex''})} & \ConcatAssocRule{}  \\
    \PRegexOr{(\PRegexOr{\Regex}{\Regex'}{\Probability_1})}{\Regex''}{\Probability_2} & $\SSREquiv$ & \PRegexOr{\Regex}{(\PRegexOr{\Regex'}{\Regex''}{\frac{(1-\Probability_1)*\Probability_2}{1 - \Probability_1 * \Probability_2}})}{\Probability_1*\Probability_2} & \OrAssociativityRule{}  \\
    \PRegexOr{\Regex{}}{\RegexAlt{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\RegexAlt{}}{\Regex{}}{1-\Probability} & \OrCommutativityRule{}\\
    \RegexConcat{\Regex{}}{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}}{\Regex{}'})}{(\RegexConcat{\Regex{}}{\Regex{}''})}{\Probability} & \DistributivityLeftRule{} \\
    \RegexConcat{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})}{\Regex{}} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}'}{\Regex{}})}{(\RegexConcat{\Regex{}''}{\Regex{}})}{\Probability} & \DistributivityRightRule{} \\
    \RegexConcat{\EmptyString{}}{\Regex{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityLeftRule{} \\
    \RegexConcat{\Regex{}}{\EmptyString{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityRightRule{} \\
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{1-\Probability} & \UnrollstarLeftRule{} \\
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{{\Regex{}}}{\Probability}}{\Regex{}})}{1-\Probability} & \UnrollstarRightRule{} 
  \end{tabular}
\end{center}

@@

\subsection{Lens Costs}
Our cost metric is based on the expected amount of information required to
recover a string in one data format from the other. We use the function
$\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ to calculate the expected amount
of information required to recover a string in $\RegexAlt$ from a string in
$\Regex$, synchronized by $\Lens$. We use the function $\LEntropyOf{\Regex
  \Given \Lens, \RegexAlt}$ to calculate the expected amount of information
required to recover a string in $\Regex$ from a string in $\RegexAlt$,
synchronized by $\Lens$.
\begin{center}
  \begin{tabular}{rcl}
    $\REntropyOf{\RegexAlt \Given \IdentityLensOf{\RegexAlt}, \RegexAlt}$
    & =
    & 0\\

    $\REntropyOf{\RegexAlt \Given \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}, \Regex}$
    & =
    & \EntropyOf{\RegexAlt}\\

    $\REntropyOf{\PRegexStar{\RegexAlt}{\ProbabilityAlt} \Given \IterateLensOf{\Lens}, \PRegexStar{\Regex}{\Probability}}$
    & =
    & $\frac{\Probability}{1-\Probability}(\REntropyOf{\RegexAlt \Given \Lens, \Regex})$\\

    $\REntropyOf{\RegexAlt_1 \Concat \RegexAlt_2 \Given \ConcatLensOf{\Lens_1}{\Lens_2}, \Regex_1 \Concat \Regex_2}$
    & =
    & $\REntropyOf{\RegexAlt_1 \Given \Lens_1, \Regex_1} + \REntropyOf{\RegexAlt_2 \Given \Lens_2, \Regex_2}$\\

    $\REntropyOf{\RegexAlt_2 \Concat \RegexAlt_1 \Given \SwapLensOf{\Lens_1}{\Lens_2}, \Regex_1 \Concat \Regex_2}$
    & =
    & $\REntropyOf{\RegexAlt_2 \Given \Lens_1, \Regex_2} + \REntropyOf{\RegexAlt_1 \Given \Lens_1, \Regex_1}$\\

    $\REntropyOf{\PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt} \Given \OrLensOf{\Lens_1}{\Lens_2}, \PRegexOr{\Regex_1}{\Regex_2}{\Probability}}$
    & =
    & $\Probability*(\REntropyOf{\Regex_1 \Given \Lens_1, \RegexAlt_1} ) + (1-\Probability)*(\REntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt_2})$\\

    $\REntropyOf{\PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt} \Given \MergeLOf{\Lens_1}{\Lens_2}, \Regex}$
    & =
    & $\ProbabilityAlt*(\REntropyOf{\RegexAlt \Given \Lens_1, \Regex_1}) + (1-\ProbabilityAlt)*(\REntropyOf{\RegexAlt \Given \Lens_2, \Regex_2})$\\

    $\REntropyOf{\RegexAlt \Given \MergeROf{\Lens_1}{\Lens_2}, \PRegexOr{\Regex_1}{\Regex_2}{\Probability}}$
    & =
    & $\Probability*(\REntropyOf{\RegexAlt \Given \Lens_1, \Regex_1} - \Log_2(\Probability)) + $\\
    &
    & $(1-\Probability)*(\REntropyOf{\RegexAlt \Given \Lens_2, \Regex_2} - \Log_2(1 - \Probability))$\\

    $\REntropyOf{\Regex \Given \InvertOf{\Lens}, \RegexAlt}$
    & =
    & $\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$\\
  \end{tabular}
\end{center}
$\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ is defined symmetrically.  These
functions calculate the expected number of bits to recover one data format from
a synchronized string in the other format.
The cost of a lens, $c(\Lens \OfType \Regex \Leftrightarrow \RegexAlt) =
\LEntropyOf{\Regex \Given \Lens, \RegexAlt} + \REntropyOf{\RegexAlt \Given
  \Lens, \Regex}$ is the
sum of recovering the left format from the right, and the right from the left.
We have proven theorems demonstrating the calculated entropy corresponds to the
actual conditional entropy to recover the data.\bcp{Again, the proofs should be
  included.}

\begin{theorem}
  Let $\Lens \OfType \Regex \Leftrightarrow \RegexAlt$, where $\Lens$ does not
  include composition, $\Regex$ and $\RegexAlt$ and unambiguous, and neither
  $\Regex$ nor $\RegexAlt$ contain any empty subcomponents.
  \begin{enumerate}
  \item $\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ calculates the expected
    information content of $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt}}$,
    given $\SetOf{s \SuchThat s \in \LanguageOf{\Regex} \BooleanAnd
      \Lens.\PutROf{s}{t} = t}$
  \item $\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ calculates the expected
    information content of $\SetOf{s \SuchThat s \in \LanguageOf{\Regex}}$,
    given $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt} \BooleanAnd
      \Lens.\PutLOf{t}{s} = s}$
  \end{enumerate}
\end{theorem}

@@

As these DNF regular expressions are \emph{stochastic}, they are annotated with
probabilities to express a probability distribution, in addition to a language.
\begin{center}
  \begin{tabular}{rcl}
    $\ProbabilityOf{\PRegexStar{\DNFRegex}{\Probability}}{\String}$
    & =
    & $\Sigma_n \Sigma_{\String = \String_1 \ldots \String_n}\Probability^n*(1-\Probability)*\Pi_{i=1}^n\ProbabilityOf{\DNFRegex}{\String_i}$\\

@@

    $\EntropyOf{\SequenceOf{\String_0 \SeqSep \Atom_1 \SeqSep \ldots \SeqSep \Atom_n \SeqSep \String_n}}$
    & =
    & $\Sigma_{i = 1}^n\EntropyOf{\Atom_i}$ \\

@@

Much like we defined a syntactic means to find the expected number of bits to
recover string from a synchronized string in another format for simple symmetric
string lenses, we do the same for symmetric DNF lenses.  Unlike simple symmetric
string lenses, DNF lenses do not have composition, so the entropy can be defined
over all terms.  $\REntropyOf{\DNFRegex \Given \SDNFLens, \DNFRegexAlt}$ is the
expected number of bits required to recover a string in $\DNFRegex$ from a
synchronized string in $\DNFRegexAlt$.

\[
  \begin{array}{rl}
    & \REntropyOf{\DNFOf{\Sequence_1 \DNFSep \ldots \DNFSep \Sequence_n}\\
    & \hspace*{1.21em}\Given
      (\SDNFLensOf{(i_1,j_1,\SSQLens_1)\DNFLSep
      \ldots\DNFLSep
      (i_p,j_p,\SSQLens_p)}
      ,\ListOf{c_1;\ldots;c_q}
      ,\ListOf{d_1;\ldots;d_r})\\
    & \hspace*{1.21em},~\DNFOf{\SequenceAlt_1 \DNFSep \ldots \DNFSep \SequenceAlt_m}}\\
    = & \frac{\Sigma_{j=1}^m(H_j)}{m}\\
    & \text{where } H_j = \frac{\Sigma_{k\SuchThat j_k = j}
      \REntropyOf{\Sequence_{i_k}\Given\SequenceLens_k,\SequenceAlt_j}}{\SizeOf{\SetOf{k\SuchThat
      j_k = j}}} + log_2\SizeOf{\SetOf{k\SuchThat j_k = j}}
  \end{array}
\]

@@

\[
  \begin{array}{rl}
      & \REntropyOf{
        \SequenceOf{\String_0,\Atom_1,\ldots,\Atom_n,\String_n}\\
      & \hspace*{1em}\Given
        (\SSQLensOf{(i_1,j_1,\SAtomLens_1)\SeqLSep
        \ldots\SeqLSep
        (i_p,j_p,\SAtomLens_p)}
        ,\ListOf{(k_1,\String_1);\ldots;(k_q,\String_q)}
        ,\ListOf{(l_1,\StringAlt_1);\ldots;(l_r,\StringAlt_r)})\\
      & \hspace*{1.21em},~
        \SequenceOf{\StringAlt_0,\AtomAlt_1,\ldots,\AtomAlt_m,\StringAlt_m}}\\
      =
      & \Sigma_{x=1}^p\REntropyOf{\Atom_{i_x} \Given \AtomAlt_{i_x},\AtomLens_x} +
        \Sigma_{x=1}^q\REntropyOf{\Atom_{k_x}}
  \end{array}
\]

@@

\[\begin{array}{rl}
      & \REntropyOf{\PRegexStar{\DNFRegex}{\Probability} \Given \StarOf{\DNFRegexAlt},\StarOf{\SDNFLens}}\\
      =\\
      & 4*\REntropyOf{\DNFRegex\Given\DNFRegexAlt,\SDNFLens}
  \end{array}
\]
The expected number of bits required to recover a string in $\DNFRegexAlt$ from
a synchronized string in $\DNFRegex$, $\REntropyOf{\DNFRegexAlt \Given
  \SDNFLens, \DNFRegex}$ is defined symmetrically.

@@

\paragraph*{Compositional Synthesis} Compositional synthesis allows us to use
previously defined lenses - whether they are previously defined by users or by
synthesis. We alter \GreedySynth such that two regular expression atoms can be
mapped to each other either if they are the same regular expression and have
examples agreeing with the examples\bcp{What does ``have
examples agreeing with the examples'' mean??}, or if there is a lens which is a
composition of predefined lenses between them, where the lens alters the
examples in a desired way.

\paragraph*{\RXSearch Permissiveness Customization}\bcp{I don't follow this:}
While the termination heuristic works for many situations, there are situations
where a large number of expansions must be performed to find the correct lens.
While users could prevent incorrect lenses with a large number of examples, we
find it simpler to let the users specify that \RXSearch should be more
restrictive in which lenses it accepts. If the user's desired transformation has
a particularly high cost, the algorithm may spend a long time searching in vain
for lenses with lower cost.  By tuning \RXSearch to me more permissive in the
lenses it accepts, this search can be short-circuited.
% end synthesis

@@

We have extended Optician\bcp{this is the first mention of the word
  Optician!} with our symmetric synthesis algorithm. Users can 
write synthesis tasks alongside lens combinators, use those synthesis results in
the manual writing of new lenses, and use previously defined lenses in the
synthesis of new lenses.

In this evaluation, we aim to answer two primary questions:
\begin{enumerate}
\item Is our synthesis procedure efficient enough to be used in everyday
  development?  In particular, can it be used in place of the existing
  bijective lens synthesis algorithm integrated with Boomerang (which
  \citet{miltner+:bijective-synthesis} have argued is reasonably fast in
  practice)? 
  \item Are the heuristics we have described useful for reducing the 
  ``specification overhead'' for users?  \bcp{Maybe it would be better to
    ask {\em which} heuristics are the most useful.  Also, are we talking
    about reducing the number of examples needed?  It's not clear.  (And:
    Don't the heuristics also make it faster??)}
\end{enumerate}
